---
title: "Main Steps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PheCAP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(PheCAP)
```

Specify the paths for features and labels. Labels can be NULL if only surrogate-assisted feature extraction (SAFE) is conducted.

```{r, eval=FALSE}
data <- PhecapData(
  icd_feature = "icd_feature.csv",
  nlp_feature = "nlp_feature.csv",
  hu_feature = "hu_feature.csv",
  training_label = "training_label.csv",
  validation_label = "validation_label.csv",
  patient_index = "PatientID")
```

Specify the surrogate used for surrogate-assisted feature extraction (SAFE). The typical way is to specify a main ICD code, a main NLP CUI, as well as their combination. In some cases one may want to define surrogate through lab test. The default lower_cutoff is 1, and the default upper_cutoff is 10. Feel free to change the cutoffs based on domain knowledge.

```{r, eval=FALSE}
surrogates <- list(
  PhecapSurrogate(
    variable_names = "ICD1",
    lower_cutoff = 1, upper_cutoff = 10),
  PhecapSurrogate(
    variable_names = "NLP1",
    lower_cutoff = 1, upper_cutoff = 10),
  PhecapSurrogate(
    variable_names = c("ICD1", "NLP1"),
    lower_cutoff = 1, upper_cutoff = 10))
```

Run surrogate-assisted feature extraction (SAFE) and show result.

```{r, eval=FALSE}
feature_selected <- phecap_run_feature_extraction(
  data, surrogates)
print(feature_selected)
```

Train phenotyping model and show the fitted model, with the AUC on the training set as well as random splits

```{r, eval=FALSE}
model <- phecap_train_phenotyping_model(
  data, surrogates, feature_selected)
print(model)
```

Validate phenotyping model using validation label, and show the AUC and ROC

```{r, eval=FALSE}
validation <- phecap_validate_phenotyping_model(
  data, surrogates, model)
print(validation)
phecap_plot_roc_curves(validation)
```

Apply the model to all the patients to obtain predicted phenotype.

```{r, eval=FALSE}
phenotype <- phecap_predict_phenotype(
  data, surrogates, model)
```

